\documentclass{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb}
\usepackage[T1]{fontenc}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{array}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=blue,      
    urlcolor=blue,
    }

\title{Apps Release Regression Governance Framework (SLO/CFU)}
\author{Jansel Valentin}
\date{March 8, 2026}

\newcommand{\CFU}{\mathrm{CFU}}
\newcommand{\MAU}{\mathrm{MAU}}
\newcommand{\EB}{\mathrm{EB}}
\newcommand{\E}{\mathrm{E}}
\newcommand{\pp}{\text{pp}}
\newcommand{\dhalt}{\delta_{\text{halt}}}
\newcommand{\dhold}{\delta_{\text{hold}}}

\begin{document}

\maketitle

\begin{abstract}
This note defines a staged-rollout governance framework for mobile app releases using 
crash-free users (CFU), error-budget accounting, and minimum-exposure gates.

The intent is simple: release decisions should not depend on gut feel.
Every threshold should be defensible through one of three lenses:

\begin{itemize}[leftmargin=1.4em]
    \item \textbf{Error-budget math}: how much of the monthly budget would this regression consume if it reached 100\% rollout?
    \item \textbf{Statistical detectability}: do we have enough exposure to trust the signal?
    \item \textbf{Time to budget depletion}: if this regression continues, how quickly would it consume the monthly budget?
\end{itemize}

\end{abstract}

\section*{Lounge app baseline}

\begin{itemize}[leftmargin=1.4em]
    \item \textbf{Monthly active user ($\MAU$)}: \(2{,}000{,}000\)
    \item \textbf{Current baseline CFU}: \(\approx 99.1\%\) \\
    \emph{Context only; this is not what defines the release thresholds.}
    \item \textbf{SLO target}: \(S = 99\%\)
    \item \textbf{SLO window}: \(W = 30\) days
    \item \textbf{Release gate window}: \(w = 6\) hours
\end{itemize}

\section{Core definitions}

Let $U$ be the number of unique users active in a window and $C$ the number of unique users who experienced at least one crash in that same window.

\paragraph{Crash-free users CFU.}

\[
\CFU = 1 - \frac{C}{U}
\]

\paragraph{Crash rate / error rate.}
\[
p = 1 - \CFU
\]

\paragraph{Allowed error rate from the SLO.}
For an SLO target $\S$
\[
\E = 1 - S
\]

For $S = 99\%$
\[
\E = 1\% = 0.01
\]

\paragraph{Monthly error budget in crashed users.}
\[
\EB = \E \cdot \MAU
\]

\paragraph{Remaining monthly error budget in crashed users.}
If the current CFU over the SLO window is $\CFU_{\mathrm{current}}$, then
\[
\EB_{\mathrm{rem}} = 
\max(0,\ \E - (1 - \CFU_{\mathrm{current}})) \cdot \MAU
\]

\paragraph{Release-induced regression.} 
Let $p_{\mathrm{new}}$ be the crash rate of the new version and $p_{\mathrm{control}}$ the crash rate of the control cohort (usually previous versions), measured over the \emph{same gate window}. 
Then:
\[
\delta = p_{\mathrm{new}} - p_{\mathrm{control}}
\]

A positive $\delta$ means the release is worse than control.
Throughout this note, $\delta$ is expressed in \emph{percentage points} $(\pp)$, not percent.

\section{Governance principle}

\paragraph{Governance rule.}
A single release should not be allowed to consume more than a fixed fraction $\alpha$ of the \emph{total monthly error budget} if it were to reach 100\% rollout and the observed regression persisted.

We choose:
\[
\alpha = 20\%
\]

\paragraph{Projected full-rollout budget share.}
If a release shows an incremental regression $\delta$, then at 100\% rollout its projected monthly excess crashed users would be:
\[
\delta \cdot \MAU
\]

As a fraction of the monthly error budget: 
\[
\text{BudgetShare}
=
\frac{\delta \cdot \MAU}{\E \cdot \MAU}
=
\frac{\delta}{\E}
\]

Also:
\[
\delta = \text{BudgetShare} \cdot \E
\]

Therefore the maximum release-induced regression allowed by the \textbf{governance rule} is:
\[
\delta_{\max} = \alpha \cdot \E
\]

From which it follows:
\[
\text{BudgetShare} < \alpha
\]

For $S = 99\%$, $\E = 1\%$, so:
\[
\delta_{\max} = 20\% \cdot \E = 0.20\pp
\]

This is the core \textbf{policy ceiling}:
\[
\delta \ge 0.20\pp \quad \Rightarrow \quad \text{release is not acceptable}
\]

\paragraph{Important clarification.}
The ceiling $0.20\pp$ is derived from:
\begin{itemize}[leftmargin=1.4em]
    \item the SLO target $\S$, through $\E = 1 - \S$, and
    \item the governance choice $\alpha$.
\end{itemize}

so:

\begin{itemize}[leftmargin=1.4em]
    \item changing $\MAU$ \textbf{does not} change the threshold in $\pp$,
    \item changing the SLO \textbf{does} change the threshold in $\pp$,
    \item changing $\alpha$ \textbf{does} change the threshold in $\pp$.
\end{itemize}

\section{Intuition for $\delta$.}

$\delta$ is the \emph{excess crash rate} of the new version relative to control.

Example:
\[
\delta = 0.10\pp = 0.001
\]

If the new version has $N = 20{,}000$ users in the gate window, the expected excess crashed users relative to control is:
\[
\delta \cdot N = 0.001 \cdot 20{,}000 = 20
\]

So a regression of $0.10\pp$ means about 20 additional crashed users per 20K exposed users, compared with control.

\section{Why exposure gates are needed}

A threshold is useless if the sample is too small to trust the signal. 

Early rollout stages have lower exposure, so they can only reliably catch \emph{larger} regressions. 
Later stages accumulate enough exposure to enforce \emph{stricter} thresholds.

This is why rollout governance needs both:
\begin{itemize}[leftmargin=1.4em]
    \item a \textbf{policy ceiling} derived from error-budget math, and
    \item an \textbf{exposure gate} derived from detectability.
\end{itemize}

\subsection{Detectability rule of thumb}

For planning purposes, use the approximate 95\% confidence margin of error for a proportion:
\[
m \approx 1.96 \sqrt{\frac{p(1-p)}{N}}
\qquad \Longleftrightarrow \qquad
N \approx \frac{1.96^2 p(1-p)}{m^2}
\]

In this framework:
\begin{itemize}[leftmargin=1.4em]
    \item use $p \approx E$, crashed users rate, i.e. near the SLO floor;
    \item use $m$ equal to the stage's $\dhalt$.
\end{itemize}

For $S = 99\%$, we use $p \approx 0.01$ as the planning point.

\paragraph{Why use $m = \dhalt$?}
Because $N_{min}$ is meant to answer a practical question:

\begin{quote}
``How many exposed users do we need before a regression large enough to justify \textbf{HALT} is distinguishable from normal sampling noise?''
\end{quote}

So we use the halt threshold itself as the planning margin:
\[
m = \dhalt
\]

This should be read as a \emph{detectability heuristic}, not as a formal proof. 
It means we size the exposure gate so that a regression at the HALT boundary is large enough to act on with reasonable confidence.

For example, if
\[
\dhalt = 0.20\pp = 0.002
\]
then over $N = 10{,}000$ exposed users, that corresponds to:
\[
0.002 \cdot 10{,}000 = 20
\]
excess crashed users relative to control.

So a halt threshold of $0.20\pp$ can be interpreted operationally as:
\begin{quote}
``At this exposure, a HALT-level regression means about 20 additional crashed users per 10K exposed users.''
\end{quote}

\paragraph{Why use $p \approx \E = 1\%$?}
Because $1\%$ is the error-rate boundary implied by the SLO:
\[
S = 99\%
\qquad \Longrightarrow \qquad
\E = 1 - S = 1\%
\]

The purpose of $p$ in the sample-size formula is not to assert the app is ``actually'' running at exactly $1\%$ crash rate.
It is to choose a \emph{planning point} near the operational boundary that matters most:
the SLO floor.

In other words, we ask:
\begin{quote}
``How much exposure would we want if the app were performing near the SLO boundary and we wanted to detect a HALT-sized regression?''
\end{quote}

That gives a conservative and policy-aligned exposure gate.

\paragraph{Example: same $\delta$, same incremental harm per $N$}
A key property of $\delta$ is that, for a fixed exposure $N$, it always maps to the same number of \emph{additional} crashed users, regardless of the starting baseline.

Let $N = 10{,}000$ and let
\[
\delta = 0.20\pp = 0.002
\]

\textbf{Case 1: control CFU $= 99.0\%$} \\
Then control crash rate is $p_{\text{control}} = 1.0\% = 0.01$:
\[
0.01 \cdot 10{,}000 = 100 \text{ crashed users }
\]
If the new version regresses by $0.20\pp$, then:
\[
p_{\text{new}} = 1.2\% = 0.012
\]
\[
0.012 \cdot 10{,}000 = 120 \text{ crashed users }
\]
Difference:
\[
120 - 100 = 20
\]

\textbf{Case 2: control CFU $= 99.5\%$}\\
Then control crash rate is $p_{\text{control}} = 0.5\% = 0.005$:
\[
0.005 \cdot 10{,}000 = 50 \text{ crashed users}
\]
If the version regresses by the same $0.20\pp$, then:
\[
p_{\text{new}} = 0.7\% = 0.007
\]
\[
0.007 \cdot 10{,}000 = 70 \text{ crashed users}
\]
Difference:
\[
70 - 50 = 20
\]

So the operational meaning of $\delta$ is stable:
\begin{quote}
For every 10K exposed users, a regression of $0.20\pp$ means about 20 extra crashed users relative to control, regardless of whether the baseline CFU was $99.0\%$ or $99.5\%$. 
\end{quote}

\paragraph{Example: why $p = 1\%$ is a conservative planning point}
Using the approximate 95\% margin-of-error formula
\[
N \approx \frac{1.96^2 p(1-p)}{m^2}
\]
with
\[
m = 0.20\pp = 0.002
\]
gives:

\textbf{If $p = 1.0\%$}
\[
N \approx \frac{1.96^2 \cdot 0.01 \cdot 0.99}{0.002^2}
\approx 9{,}508
\]
which rounds to about $10K$.

\text{If $p$ were chosen $p = 0.5\%$:}
\[
N \approx \frac{1.96^2 \cdot 0.005 \cdot 0.995}{0.002^2}
\approx 4{,}776
\]

So using $p = 1\%$ requires more exposure than using $p = 0.5\%$.
That is why $p = \E = 1\%$ is a reasonable planning choice: it anchors the gate near the SLO boundary and avoids underestimating the exposure needed to trust a HALT-level signal.

\paragraph{Operational meaning of $N_{\min}$.}
$N_{\min}$ is a \emph{practical gating heuristic}, not a formal proof that the release is safe.
It answers:
\begin{quote}
``Do we have enough new-version exposure to trust a signal large enough to justify \textbf{HALT}?''
\end{quote}

It is intentionally tied to $\dhalt$, not to $\dhold$. 
That is why the hold band can be stricter than what $N_{\min}$ alone fully resolves:
\textbf{HOLD} is a softer precautionary action than \textbf{HALT}.


\section{Decision framework}

We define three actions:

\begin{itemize}[leftmargin=1.4em]
    \item \textbf{HALT}: stop rollout progression immediately.
    \item \textbf{HOLD}: remain at the current rollout state and continue observing; assign an owner to issues.
    \item \textbf{RAMP}: continue to the next rollout stage.
\end{itemize}

The design intent is:

\begin{itemize}[leftmargin=1.4em]
    \item early stages mainly catch \emph{large} regresions,
    \item later stages enforce \emph{tighter} regression limits,
    \item no stage may tolerate a regression that violates the global budget ceiling.
\end{itemize}

\paragraph{Hard requirements for the thresholds.}
For each stage:
\begin{enumerate}[leftmargin=1.4em]
    \item $\dhalt$ must not exceed the global ceiling implied by the \textbf{governance rule}.
    \item $N_{\min}$ must be feasible within the observation window for that stage under normal traffic.
\end{enumerate}

\section{Decision matrix}

\begin{table}[h]
    \centering
    \small
    \renewcommand{\arraystretch}{1.2}
    \begin{tabular}{@{}r c c c@{}}
        \toprule
        \textbf{Rollout stage} &
        \textbf{\(N_{\min}\) (new-version users in last 6h)} &
        \textbf{\(\delta_{\text{hold}}\)} &
        \textbf{\(\delta_{\text{halt}}\)} \\
        \midrule
        1\%   & 10K  & 0.10\pp & 0.20\pp \\
        2\%   & 10K  & 0.10\pp & 0.20\pp \\
        5\%   & 10K  & 0.10\pp & 0.20\pp \\
        10\%  & 20K  & 0.10\pp & 0.20\pp \\
        20\%  & 50K  & 0.05\pp & 0.10\pp \\
        50\%  & 100K & 0.05\pp & 0.10\pp \\
        100\% & 100K & 0.05\pp & 0.10\pp \\
        \bottomrule
    \end{tabular}

    \vspace{0.3em}
    \parbox{0.92\linewidth}{\footnotesize
    \textit{Decision rule once \(N_{\min}\) is met:}
    RAMP if \(\delta < \delta_{\text{hold}}\);
    HOLD if \(\delta_{\text{hold}} \le \delta < \delta_{\text{halt}}\);
    HALT if \(\delta \ge \delta_{\text{halt}}\).
    At 100\% rollout, ``RAMP'' means ``continue rollout / do not trigger rollback based on this gate''.
    }
\end{table}

\paragraph{Interpretation.}
The table intentionally uses:
\begin{itemize}[leftmargin=1.4em]
    \item \textbf{looser thresholds} at low rollout, where exposure is limited;
    \item \textbf{stricter thresholds} at higher rollout, where both confidence and potential customer har are greater. 
\end{itemize}

\paragraph{Note on $N_{\min}$.}
$N_{\min}$ is chosen to make $\dhalt$-level signals credible enough to act on.
It is \emph{not} claiming equal statistical resolution for $\dhold$.

\section{Worked example}

Suppose:
\begin{itemize}[leftmargin=1.4em]
    \item rollout stage $= 5\%$,
    \item new-version users in last 6h $= 15K$,
    \item observed regression $= 0.04\pp$.s
\end{itemize}

From the table:

\begin{itemize}[leftmargin=1.4em]
    \item $N_{\min} = 10K$ \quad (met),
    \item $\dhold = 0.10\pp$,
    \item $\dhalt = 0.20\pp$.
\end{itemize}

Since
\[
0.04\pp < 0.10\pp
\]
the decision is:
\[
\textbf{RAMP}
\]

\section{What changes when app scale changes?}

\subsection{Changing $\MAU$}

A larger $\MAU$ does \emph{not} change the thresholds expressed in $\pp$. 
It only changes how quickly the app can accumulate the required exposure in a fixed gate
window.

So if another app has much higher traffic, it may be able to:
\begin{itemize}[leftmargin=1.4em]
    \item keep the same governance thresholds, but
    \item shorten the gate window from 6h to 3h or 2h because it reaches $N_{\min}$
    faster.
\end{itemize}

This is an \emph{exposure-speed} difference, not a policy difference.

\subsection{Changing governance strictness}

Is a different app chooses a stricter rule such as:
\[
\alpha = 10\%
\]
then the global ceiling becomes:
\[
\delta_{\max} = 0.10 \cdot \E
\]

For $S = 99\%$, that would mean:
\[
\delta_{\max} = 0.10\pp
\]

That is a \textbf{governance choice}, not something forced by higher traffic.

\subsection{Changing the SLO}

If the SLO changes, the thresholds in $\pp$ must be recomputed because:
\[
\delta_{\max} = \alpha \cdot (1 - S)
\]

Example:
\begin{itemize}[leftmargin=1.4em]
    \item if $S = 99\%$, then $\delta_{\max} = 0.20\pp$,
    \item if $S = 99.5\%$, then $\delta_{\max} = 0.10\pp$,
\end{itemize}
assuming the same $\alpha = 20\%$.

So the table is portable across apps with different $\MAU$, but \emph{not automatically across different SLO targets}

\section{Release policy}

At each rollout stage, using the 6h gate window:

\begin{enumerate}[leftmargin=1.8em]
    \item If blocker crash affects a critical journey (for example strtup, login, checkout, cart) with impact $\ge 5$ users, assign owner and investigate immediately $\rightarrow$ \textbf{HALT}. Incident / hotfix process follows separately.
    \item If $\texttt{users\_new\_6h} < N_{\min}(\text{stage})$, $\rightarrow$ \textbf{HOLD}.\\
    No decision to ramp is allowed before the exposure gate is met.
    \item Compute $\delta$ against control over \emph{same} 6h window.
    \item If $\delta \ge \dhalt$, $\rightarrow$ \textbf{HALT}.
    \item Else if $\delta \ge \dhold$, $\rightarrow$ \textbf{HOLD}.
    \item Else $\rightarrow$ \textbf{RAMP}.
\end{enumerate}

\section{Low remaining budget}

If the remaining monthly budget is already low, the normal thresholds may be too permissive.

The mathematically correct ceiling is:
\[
\delta_{\max,\mathrm{effective}} = \min\left(\alpha \cdot \E, \frac{\EB_{\mathrm{rem}}}{\MAU}\right)
\]

This says:
\begin{quote}
the release must not be allowed to project beyond either the normal per-release budget cap or the budget that is actually left.
\end{quote}

\paragraph{Simple operational rule.}
As a practical simplication, if
\[
\EB_{\mathrm{rem}} < 20\% \cdot \EB
\]
then tighten all stages to:
\[
\dhold = 0.05\pp,
\qquad
\dhalt = 0.10\pp
\]

This simplification is conservative, but it is still only a simplification. 
If the remaining budget is extremely low, the dynamic ceiling above should take precedence.

\section{Observation window}

The recommended enforcement window is 6 hours because it is practical compromise between:
\begin{itemize}[leftmargin=1.8em]
    \item catching regressions early enough to limit harm, and
    \item giving the release enough time to accumulate the exposure required by the gate.
\end{itemize}

Apps with higher traffic may shorten this window if they still hit $N_{\min}$ reliably.
Apps with lower trafic may need a longer gate window.

\section*{Appendix: Burn rate framing}

Shows how \emph{incremental regression} relates to \emph{burn rate}

\subsection*{Classic burn rate}
Allowed error rate: $E = 1 - S$.

Observed error rate in a window: $\delta$.

Burn rate:
\[
\mathrm{BR} = \frac{\delta}{e}
\]

\subsection*{Release regression burn rate (full rollout)}

\[
\mathrm{BR}_{\text{reg,100\%}} = \frac{\delta}{e}
\]

This is the same quantity as $\texttt{BudgetShare}$ above.

\subsection*{Stage-aware burn rate (what you use during rollout)}
At rollout stage, only a fraction $f$ (stage \%) of traffic is on the new version. The current incremental burn rate on the whole population is:
\[
\mathrm{BR}_{\text{stage}} = \frac{f \cdot \delta}{e}
\]

Example: $f=0.1$ (10\% rollout) and $\delta/e=0.1$:
\[
\mathrm{BR}_{\text{stage}} = 0.1 \cdot 0.1 = 0.01
\]
Meaning: at the current stage the release burns \textbf{1\% of the monthly budget pace} (incrementally).

\subsection*{Time to exhaust remaining budget (optional narrative)}
Let $r$ be the \emph{remaining budget fraction} (e.g.\ 10\% remaining $\Rightarrow r=0.10$). If the regression persists:
\[
T_{\text{exhaust}} \approx \frac{r \cdot 30\text{d}}{\mathrm{BR}_{\text{stage}}}
\]

This supports the narrative:
\begin{quote}
At current regression $\delta$ and rollout fraction $f$, we're burning at $\mathrm{BR}_{\text{stage}}$. If left unchanged, we'll consume the remaining $r$ budget in approximately $T_{\text{exhaust}}$.
\end{quote}

\section*{Appendix: References}

\begin{itemize}[leftmargin=1.4em]
    \item Defining SLOs: \href{Defining SLOs}{https://sre.google/workbook/implementing-slos/}
    \item Alerting on SLOs: \href{Alerting on SLOs}{https://sre.google/workbook/alerting-on-slos/}
\end{itemize}

\end{document}
